---
title: "Exploratory Analysis"
author: "Ying Jin"
date: "2020/10/1"
output: github_document
---


```{r setup}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot_cpntinuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Load Weather dataset

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## Introduce `group_by`

```{r}
weather_df %>% 
  group_by(name,month) %>% 
  ungroup(name)
```


## Counting things

count month / month observations

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarise(n_obs = n())
```

we can use "count"
 
```{r}
weather_df %>% 
  count(name,month,name = "n_obs")
```

**NEVER** use base R's table

```{r,eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
```

other helpful counters

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarise(n_obs = n(),
            n_days = n_distinct(date))
  
```

## A diggression on 2 x 2 tables

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~"cold",
      tmax >= 5 ~"not_cold",
      TRUE ~""
    )
  ) %>% 
  group_by(name,cold) %>% 
  summarise(count = n())
```

A better one to create 2 x 2 table

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~"cold",
      tmax >= 5 ~"not_cold",
      TRUE ~""
    )
  ) %>% 
  janitor::tabyl(name,cold)
```

## General summaries

You can do lots of summaries

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarise(
    mean_tmax = mean(tmax,na.rm = TRUE),
    mean_prcp = mean(prcp,na.rm = TRUE),
    median_tmin = median(tmin,na.rm = TRUE)
  )
```

This is a dataframe! So we can make a plot.

```{r}
weather_df %>% 
  group_by(name,month) %>% 
  summarise(
    mean_tmax = mean(tmax,na.rm = TRUE),
    mean_prcp = mean(prcp,na.rm = TRUE),
    median_tmin = median(tmin,na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax,color = name)) +
  geom_point() +
  geom_line()
```

Suppose you want to summarise many columns.

```{r}
 weather_df %>% 
  group_by(name,month) %>% 
  summarise(across(prcp:tmin,mean))
```

Reminder: sometimes your results are easier to read in another format.

```{r}
weather_df %>% 
  group_by(name,month) %>% 
  summarise(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 1)
```

## group_by and mutate

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax,na.rm =TRUE),
    centered_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date,y = centered_tmax,color = name)) +
  geom_point()
```

What about window functions?

ranking

```{r}
weather_df %>% 
  group_by(name,month) %>% 
  mutate(
   temp_rank = min_rank(tmax)
  ) %>% 
  filter(temp_rank == 1)
```

lag

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(lag_temp = lag(tmax),
         temp_change = tmax - lag_temp) %>% 
  summarise(
    temp_change_max = max(temp_change,na.rm = TRUE),
    temp_change_sd = sd(temp_change,na.rm = TRUE)
  )
```

## Quick note

Summarize only gets you so far
